### 2024-03-12 output/paper/paper.qmd
change

---
title: "Untitled"
author: "Yunshu Zhang"

thanks: "Code and data are available at: https://github.com/Yunshu921/mortality_in_Alberta.git."
format: pdf
editor: visual
date: today
date-format: long
number-sections: true
toc: TRUE
---

# Introduction

# Data

# Results

# Discussion

# Conclusion

\newpage

# References